---
title: "KvalReg: scale"
subtitle: "Rasch analysis"
title-block-banner: "#870052"
title-block-banner-color: "#FEEEEB"
author: 
  name: Magnus Johansson
  affiliation: Karolinska Institutet
  orcid: 0000-0003-1669-592X
  email: magnus.johansson.3@ki.se
date: last-modified
date-format: iso
always_allow_html: true
format: 
  html:
    toc: true
    toc-depth: 3
    toc-title: "Table of contents"
    embed-resources: true
    standalone: true
    page-layout: full
    code-overflow: wrap
    code-fold: true
    code-tools: true
    code-link: true
    number-sections: true
    fig-dpi: 96
    layout-align: left
    linestretch: 1.6
    theme:
      - materia
    css: styles.css
    license: CC BY
execute:
  echo: true
  warning: false
  message: false
  cache: true
editor_options: 
  markdown: 
    wrap: 72
  chunk_output_type: console
---

```{r}
library(tidyverse)
library(janitor)
library(car)

cores <- 4
library(easyRasch)
options(rgl.useNULL = TRUE)

### some commands exist in multiple packages, here we define preferred ones that are frequently used
select <- dplyr::select
count <- dplyr::count
recode <- car::recode
rename <- dplyr::rename
```

## Data import

```{r}
# read datafile, preprocessed in script_260303.R
d_all <- read_csv("data/cisiAnalysisSample.csv") %>% 
  na.omit() # may not be needed or useful?

# a vector of ordered response category labels
resp_cats <- c("Normal", "Slowness and/or minimal cognitive problems", "Mild cognitive problems, no limitations", "Moderate cognitive problems, help is required for some basic ADL", "Severe cognitive problems, help is required for most or all basic ADL", "Mild to moderate cognitive problems, does not need help for basic ADL", "Severely disabled, helpless, complete assistance needed")

# Here it would be nice to have the item content
items <- c()

# the itemlabels object is useful later on
#itemlabels <- data.frame(itemnr = names(data), item = items)
```

```{r}
# subset items to their own dataframe
d_dif <- d_all
d <- d_all %>% 
  select(motorSigns,disability,motorComplications,cognition)
itemlabels <- data.frame(itemnr = names(d), item = names(d))
```


## Figures

::: panel-tabset
### Guttman heatmap
```{r}
RIheatmap(d)
```

### Stacked bars
```{r}
RIbarstack(d)
```

### Tileplot n
```{r}
RItileplot(d_all[,-1])
```

### Tileplot %
```{r}
RItileplot(d_all[,-1], percent = T)
```
### Columns
```{r}
RIitemcols(d)
```
:::


## Rasch 1

::: panel-tabset
### Item-restscore bootstrap
```{r}
RIbootRestscore(d, iterations = 250, samplesize = 800, cpu = cores)
```
### Item fit
```{r}
simfit1 <- RIgetfit(d, 400, cores)
RIitemfit(d, simfit1)
```
### Item-restscore
```{r}
RIrestscore(d)
```
### CICC
```{r}
RIciccPlot(d)
library(RASCHplot) # devtools::install_github("ERRTG/RASCHplot")

CICCplot(PCM(d),
         which.item = c(1:4),
         lower.groups = c(0,5,10,15,20), # may need adjustment
         grid.items = TRUE)
```

### PCA of residuals
```{r}
#| tbl-cap: "PCA of Rasch model residuals"
# this takes some time...
RIbootPCA(d, iterations = 250, cpu = cores)$p99 
RIpcmPCA(d) 
```

### Residual correlations
```{r}
RIpartgamLD(d)
simcor1 <- RIgetResidCor(d, iterations = 500, cores)
RIresidcorr(d, cutoff = simcor1$p99)
```

### Many items ICC
```{r}
#RIitemCats(d)
mirt(d, model=1, itemtype='Rasch', verbose = FALSE) %>% 
  plot(type="trace", as.table = TRUE, 
       theta_lim = c(-5,6)) # changes x axis limits
```

### Targeting
```{r}
#| fig-height: 6
# increase fig-height above as needed, if you have many items
RItargeting(d, fast_thetas = T)
```
### Reliability
```{r}
RIreliability(d)
```
### DIF sex
```{r}
# always check tileplots for empty cells or low counts (below 4)
RIdifTileplot(d, factor(d_dif$sex))
RIdifTableLR(d, factor(d_dif$sex))
RIdifThreshTblLR(d, factor(d_dif$sex))
RIdifThreshFigLR(d, factor(d_dif$sex))
RIpartgamDIF(d, factor(d_dif$sex))

RIciccPlot(d, dif = "yes", dif_var = factor(d_dif$sex))
```
### DIF age
```{r}
# tree model to determine cut points
RIdifTable(d, d_dif$age)
dif_age_cut <- cut(d_dif$age, breaks = c(0,73,79,86,120))
# check tileplots for empty cells
RIdifTileplot(d, dif_age_cut)
RIdifFigure(d, d_dif$age)
RIdifFigThresh(d, d_dif$age)

# we can also use the LR and CICC method with the cutpoints
RIciccPlot(d, dif = "yes", dif_var = dif_age_cut)
RIdifTableLR(d, dif_age_cut)
RIdifFigureLR(d, dif_age_cut)
RIdifThreshTblLR(d, dif_age_cut)
RIdifThreshFigLR(d, dif_age_cut)
```

wrangling code if needed
```{r}
#| eval: false

d %>% 
  mutate(across(everything(), ~ car::recode(.x,"6=5"))) %>% 
  RIdifTable(d_dif$age)
```
### age+gender
```{r}
# interaction DIF?
RIdifTable2(d, 
              dif.var1 = d_dif$age,
              dif.var2 = factor(d_dif$sex))
```
### Person fit
```{r}
RIpfit(d)
RIu3poly(d) # this only produces a stable % of misfits

# who is misfit according to U3?
params <- RIitemparams(d, output = "dataframe") %>%
  select(!Location) %>%
  as.matrix()
thetas <- RIestThetas(d)$WLE

cf <- PerFit::U3poly(matrix = d, Ncat = max(as.matrix(d)) +
  1, IRT.PModel = "PCM", IP = params, Ability = thetas)

flagged <- PerFit::flagged.resp(cf)
flag_id <- flagged[["Scores"]] %>% 
  as.data.frame() %>% 
  pull(FlaggedID)
```
### Item-restscore bootstrap nomisfits
```{r}
RIbootRestscore(d[-flag_id,])
```
### Item fit nomisfits
```{r}
simfitx <- RIgetfit(d[-flag_id,], iterations = 400, cpu = cores)
RIitemfit(d[-flag_id,], simfitx)
```
### LD nomisfits
```{r}
RIpartgamLD(d[-flag_id,])
simcor1b <- RIgetResidCor(d[-flag_id,], iterations = 500, cores)
RIresidcorr(d[-flag_id,], cutoff = simcor1b$p99)
```
### Reliability
```{r}
RIreliability(d) # for PSI and RMU
RIrelRep(d) # alpha and omega
RIrelRep(d, rel = "omega")
```
### Item parameters
```{r}
RIitemparams(d)
# change the filename as needed
RIitemparams(d, output = "file", filename = "scale_parameters.csv")
```
### Transformation table
```{r}
RIscoreSE(d)
# change the filename as needed
RIscoreSE(d, output = "dataframe") %>% 
  write_csv("scale_transformation.csv")
```
### Distribution of thetas

Using Weighted Likelihood Estimation (Warm, 1989).

```{r}
hist(RIestThetas(d)$WLE, breaks = 30, col = "lightblue", main = "Histogram of WLE thetas", xlab = "Latent logit scale")
summary(RIestThetas(d)$WLE)
```
:::

## CFA

```{r}
library(lavaan)
library(lavaanExtra)
library(lavaanPlot)
library(dynamic) # remotes::install_github("melissagwolf/dynamic")
# see https://rpubs.com/dmcneish/1025400 for details

fit_metrics_scaled <- c("chisq.scaled", "df", "pvalue.scaled",
                        "cfi.scaled", "tli.scaled", "rmsea.scaled",
                        "rmsea.ci.lower.scaled","rmsea.ci.upper.scaled",
                        "srmr")
```

```{r}
latent <- list(f1 = names(d)) # use names of all variables to load on one variable
model <- write_lavaan(latent = latent) # create the model specification
# fit the CFA and store result as cfa1
cfa1 <- cfa(model, data = d,
            ordered = TRUE, estimator = "WLSMV")
# review summary, including standardized factor loadings "Std.lv"
#summary(cfa1, standardized = T)

# get dynamic cutoffs
# WARNING: takes some time to run...
dyn1 <- dynamic::catOne(cfa1, reps = 500)
dyn1 # note that empirical values are disordered, CFI and RMSEA should change place
```


Review observed/emprical fit measures

```{r}
fitmeasures(cfa1, fit_metrics_scaled) %>%
  rbind() %>%
  as.data.frame() %>%
  mutate(across(where(is.numeric),~ round(.x, 3))) %>%
  knitr::kable()
```

### Modification indices

Residual correlations are indicated by op = ~~

```{r}
modificationIndices(cfa1,
                    standardized = T) %>%
  as.data.frame(row.names = NULL) %>%
  filter(mi > 3) %>%
  arrange(desc(mi)) %>%
  mutate(across(where(is.numeric),~ round(.x, 3))) %>%
  knitr::kable()
```

### Standardized loadings plot

```{r}
# https://labasim.github.io/lavaanPlot/articles/Intro_to_lavaanPlot.html
lavaanPlot(cfa1, 
           node_options = list(shape = "box", fontname = "Helvetica"),
           edge_options = list(color = "grey"), 
           graph_options = list(rankdir = "LR"), # orient plot Left to Right (LR)
           coefs = TRUE, stand = TRUE)

```


## System information
```{r}
sessionInfo()
```

